\chapter{Inverted Lists Compression For SIREn}{%
The SIREn indexing model generates a specific distribution of values. Since
each stream of values are local to an other (i.e attributes identifiers are
local to an entity, values identifiers are local to an attribute, and positions
are local to a value), a new compression algorithm that is better adapted to
the SIREn use case was researched.

The performance aspects looked after are
\begin{inparaenum}[(1)]
\item the compression
\item the decompression and
\item the compression ratio.
\end{inparaenum}
The decompression speed is important when processing queries because it means
that less time is to be spent on IO operations. The compression ratio is
important as per the storage space, but not only. As previously, less data
read from disk means less time wasted on IO operations. The compression speed
aspect does not matter when processing queries, but takes an important place
when updating the index. Indeed SIREn possesses an incremental index update
policy, thus compression algorithm showing good decompression speed will allow
indexes to be updated faster than an algorithm with slower compression
algorithm.

In this chapter, two of the compression techniques researched on are presented,
with \emph{RiceFOR} first, an algorithm that fuses together the Rice and FOR
algorithms, and the \emph{Adaptive Frame Of Reference} (AFOR), a new algorithm
based on FOR, which goal is to decrease the FOR sensibility to outliers.
}%
\label{chap:cmp-methods}
\input{Methods/compression}

\chapter{SkipBlock: Self-Indexing for Block-Based Inverted Lists}{%
SkipBlock is a model for block-based inverted lists, that can be applied as
well on traditional inverted lists. It is based on an adaptation of the Skip
List data structure. This work is orthogonal to related work on self-indexing
structures, since they might be extended to this model.

In this section, the SkipBlock model is presented with its associated search
algorithm. We discuss then the impact of the structure's parameters for both
Skip List and SkipBLock. We argue after the different strategies available to
search within intervals for both structures. Finally we compare thanks to a
cost-based model the differences between the Skip Lists and the SkipBlock. }
\label{chap:skipblock}
\input{Methods/self-indexing}