\chapter{Introduction}{
The Web as we know it is mostly a huge collection of documents, e.g. texts,
pictures, \ldots, connected together thanks to HTTP links.
These links carry no meaning about the
relationship between two documents, or about the content a document exhibits.
Linked Data refers to those same documents, but enriched with \emph{semantic}
meta-data. For example, the content of a document can be described by this
meta-data by indicating the language it was written in, or the content's type,
e.g. book or article. Tim Berners-Lee has first introduced the idea of such
enriched web documents, called \emph{Linked Data}. For such a vision to be
viable as the web is now a large collection of heterogeneous data, a framework
that sets the best practices for publishing and searching Linked Data is a
necessity. The \emph{Semantic Web} is such a framework and is not aimed to be a
replacement of the Web as we know it, but as an extension giving relevant
knowledge about documents on the web.
\linebreak

Over the past years, the quantity of Linked Data available has been increasing
and is now reaching a climax. Semantic Data now at hand, the challenge is to be
able to search for information efficiently. Linked Data is commonly stored
within databases called \emph{Triple Stores}, where querying is normally done
thanks to the standardized query language SPARQL defined with the Semantic Web.
However this language, while being very expressive, is known not to scale
efficiently to a large amount of data. \emph{SIREn} is an Information
Retrieval-based search engine which goal is to store and to query efficiently
at the web scale Linked Data. SIREn is a low level application, meant to be
used by other applications such as web services that provide people
information results about what they queried for. \emph{Sindice} is a web
service that provide people and machines a search engine over Linked Data
through the use of SIREn.
\linebreak

SIREn builds an index over the Linked Data collections for searching and
retrieval purpose. With the amount of Linked Data growing quickly, the size of
such indexes and the performance to retrieve data from it are then points of
interest to build a scalable system. A compressed index is of course important
in order to take less space, however decreasing the time spent on IO operations
is also important in order to increase the query throughput. Within the long
studied domain of Information Retrieval, compression methods have been
developed for that purpose. However they were built for traditional indexes,
i.e., indexes built on collections with textual documents (unstructured data). On
the contrary, SIREn builds indexes on Linked Data, i.e., (semi) structured
data. Because the type of data differs, the model used to index documents
is different, which implies that the distribution of indexed values are
different. State of the art compression techniques are not adapted to this new
distribution of values, and so a new compression class has been developed to
match this need: \emph{Adaptive Frame Of Reference}.
\linebreak

Index compression is one aspect that has to be taken care of when building an
efficient and scalable Information Retrieval search engine. Some information
in the index is not relevant to a query. Reading unnecessary data when
processing a query is then wasted time. An other aspect to optimize the
performance of a search engine is to avoid reading useless information with
regards to a query as much as possible by using \emph{self-indexing}
techniques. The \emph{Skip Lists} is a data structure that can be used for that
purpose. Highly efficient compression techniques are nowadays block-based
algorithms. However Skip Lists abstracts from the block view, discarding
statistics about the block that could be used for a more efficient structure.
\emph{SkipBlock} is a block-based self-indexing model that generalizes the
Skip Lists structure.

We review in this report some techniques that aim to reduce the IO access time
overhead in order to increase the overall performance of the Information
Retrieval web search engine. We present AFOR, a compression method that
provides both fast compression and decompression speed, and yet with a high
compression ratio. We propose SkipBlock, a new self-indexing model that
reduces the time of random lookups from inverted lists. }
\input{prelude/introduction}

\chapter{Digital Enterprise Research Institute} {
DERI is a worldwide organisation of research institutes with the common
objective of integrating semantics into computer science and understanding how
semantics can improve computer engineering in order to develop information
systems collaborating on a global scale. A major step in this project is the
realisation of the Semantic Web.
}
\input{prelude/DERI}